Supporting Information\\ \textit{Cross Validation Approaches for Penalized Cox Regression}

### 2024-01-12 main.tex
update figure 1 and cv deviance residual notation

The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross-validated error, analogous to using the residual sum of squares in linear regression. We refer to this method as the \emph{cross-validated deviance residuals} approach.

>>>>

The sum of squared cross-validated deviance residuals, $\sum_{i}{d}^{\text{cv}}_{i}^2$, are then used as the cross-validated error, analogous to using the residual sum of squares in linear regression. We refer to this method as the \emph{cross-validated deviance residuals} approach.

### 2024-01-12 supplementary.tex
Edits to abstract and intro
update figure 1 and cv deviance residual notation

As described in the manuscript, the Martingale residuals $\{m_i\}_{i=1}^n$ and deviance residuals $\{d_i\}_{i=1}^n$ for the Cox model are given by
\begin{align*}
  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\beta) \\
  d_{i} &= \text{sign}(m_{i})\sqrt{-2(m_i + \delta_{i}\log(\delta_{i} - m_i))},
\end{align*}
where $H_0$ is the cumulative baseline hazard.  Our cross-validated versions of these residuals use
\begin{align*}
  m_{i}^{cv} &= \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}) \\
  d_{i}^{cv} &= \text{sign}(m_{i}^{cv})\sqrt{-2(m_i^{cv} + \delta_{i}\log(\delta_{i} - m_i^{cv}))},
\end{align*}
for $i$ in the test set $D_k$, with $\hat{\beta}^{-k}$ obtained from the training set $T_k$.  The sum of squared cross-validated deviance residuals, $\sum_{i}(d_{i}^{cv})^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression.  In the manuscript, this method is referred to as the \emph{cross-validated deviance residuals} approach.

@@

\begin{algorithm}
\SetAlgoLined
Use all observations' time and status to compute $\hat{H}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}$\;
\For{each regularization parameter $\lambda$}{
 \For{each fold $k = 1,2,...,K$}{
  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
  	\For{each obervation $i$ in the $k$th fold}{
  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
  	}
 }
 Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
 }
 \caption{\label{alg:na} Compute cross-validated deviance residuals based on the Nelson-Aalen estimator}
\end{algorithm}

@@

In the original proposal of the Martingale Residuals \citep{Therneau1990}, Breslow's estimator \citep{Breslow1974} was the prefered approach for estimating the cumulative baseline hazard. Under the assumption that there is no ties among the failure times, the Breslow's estimator is defined as:
\begin{equation} 
\hat{H}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}.
\end{equation}
Covariates of the fitted model are incorporated in the denominator of the summation.
Procedure~\ref{alg:brkp} includes an explicit description of this algorithm.
\begin{algorithm}
\SetAlgoLined
\For{each regularization parameter $\lambda$}{
 Fit the model using all observations' time and status \;
 Obtain linear predictor $\hat{\eta}$ from the fitted model\;
 \uIf{Use the Breslow's estimator}{
 $\hat{H}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}$;
 }\ElseIf{Use the Kalbfleisch and Prentice estimator}{
  $\hat{H}_0(t) = \sum_{t_{i} \leq t} \left\{1 -  \left(1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})} \right) ^{\exp(-\hat{\eta}_{i})}\right\}$\;
 }
 \For{each fold $k = 1,2,...,K$}{
  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
  	\For{each obervation $i$ in the $k$th fold}{
  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
  	}
 }
 Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
 }
 \caption{\label{alg:brkp} Compute cross-validated deviance residuals based on the Breslow estimator and the Kalbfleisch and Prentice estimator}
\end{algorithm}

@@

\begin{table}[h]

@@

\begin{table}[h]
\setlength{\tabcolsep}{3pt}
\caption{\label{Tab:wb2} Weibull Distribution, scale = 1, shape = 1.5}
\centering
\begin{tabular}[t]{lccccccccc}
\toprule
 Method & $\lambda$ & log(MSE Ratio) & Brier (25$\%$) & Brier(Median) & Brier(75$\%$) & KL (25$\%$)& KL (Median) & KL (75$\%$) & C Index\\
\midrule
&\multicolumn{9}{c}{\textbf{Scenario 1}: n = 150, p = 50, 5 non-zero features, weak signal}\\
\cline{2-10}
V VH  & 0.096 (0.026) & 1.502 (0.89) & 0.185 (0.006) & 0.244 (0.015) & 0.182 (0.014) & 0.558 (0.02) & 0.681 (0.032) & 0.545 (0.034) & 0.614 (0.03) \\
Standard  & 0.089 (0.03) & 1.479 (0.873) & 0.184 (0.007) & 0.243 (0.015) & 0.182 (0.013) & 0.557 (0.021) & 0.681 (0.032) & 0.544 (0.033) & 0.615 (0.031) \\
LinearPred  & 0.09 (0.027) & 1.482 (0.871) & 0.185 (0.007) & 0.243 (0.014) & 0.182 (0.014) & 0.558 (0.02) & 0.681 (0.031) & 0.544 (0.033) & 0.615 (0.03) \\
DevResid  & 0.098 (0.025) & 1.503 (0.869) & 0.185 (0.006) & 0.244 (0.015) & 0.182 (0.014) & 0.558 (0.02) & 0.681 (0.031) & 0.545 (0.033) & 0.615 (0.03) \\
\addlinespace
&\multicolumn{9}{c}{\textbf{Scenario 2}: n = 150, p = 50, 5 non-zero features, strong signal}\\
\cline{2-10}
 V VH  & 0.07 (0.011) & 1.733 (0.83) & 0.147 (0.009) & 0.176 (0.012) & 0.128 (0.009) & 0.453 (0.023) & 0.525 (0.028) & 0.395 (0.026) & 0.75 (0.013) \\ 
Standard  & 0.066 (0.011) & 1.693 (0.832) & 0.147 (0.009) & 0.176 (0.011) & 0.128 (0.009) & 0.452 (0.023) & 0.525 (0.028) & 0.395 (0.026) & 0.749 (0.014) \\
 LinearPred  & 0.067 (0.01) & 1.699 (0.834) & 0.147 (0.009) & 0.176 (0.011) & 0.128 (0.009) & 0.452 (0.023) & 0.525 (0.028) & 0.395 (0.026) & 0.749 (0.014) \\
 DevResid  & 0.096 (0.013) & 2.14 (0.799) & 0.151 (0.009) & 0.181 (0.012) & 0.13 (0.009) & 0.463 (0.022) & 0.539 (0.028) & 0.406 (0.026) & 0.751 (0.013) \\
\addlinespace
&\multicolumn{9}{c}{\textbf{Scenario 3}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
\cline{2-10}
V VH  & 0.126 (0.019) & 2.948 (0.442) & 0.188 (0.005) & 0.252 (0.011) & 0.191 (0.011) & 0.564 (0.016) & 0.696 (0.023) & 0.569 (0.026) & 0.585 (0.043) \\ 
Standard  & 0.103 (0.024) & 2.866 (0.465) & 0.184 (0.006) & 0.243 (0.014) & 0.183 (0.012) & 0.555 (0.019) & 0.68 (0.029) & 0.548 (0.03) & 0.6 (0.041) \\ 
LinearPred  & 0.104 (0.024) & 2.869 (0.464) & 0.185 (0.006) & 0.244 (0.014) & 0.183 (0.012) & 0.556 (0.019) & 0.68 (0.029) & 0.549 (0.03) & 0.6 (0.041) \\   
 DevResid  & 0.107 (0.023) & 2.88 (0.458) & 0.185 (0.006) & 0.244 (0.013) & 0.184 (0.012) & 0.556 (0.019) & 0.682 (0.028) & 0.55 (0.03) & 0.599 (0.041) \\
\addlinespace
&\multicolumn{9}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, strong signal}\\
\cline{2-10}
 V VH  & 0.082 (0.012) & 3.74 (0.554) & 0.145 (0.019) & 0.173 (0.029) & 0.127 (0.02) & 0.443 (0.05) & 0.519 (0.068) & 0.395 (0.058) & 0.766 (0.047) \\  
Standard  & 0.059 (0.009) & 3.414 (0.582) & 0.131 (0.015) & 0.154 (0.02) & 0.117 (0.014) & 0.405 (0.041) & 0.467 (0.05) & 0.367 (0.044) & 0.78 (0.03) \\   
LinearPred  & 0.06 (0.009) & 3.435 (0.583) & 0.132 (0.015) & 0.154 (0.02) & 0.117 (0.014) & 0.406 (0.041) & 0.468 (0.051) & 0.366 (0.043) & 0.78 (0.031) \\   
 DevResid  & 0.074 (0.008) & 3.66 (0.51) & 0.139 (0.014) & 0.163 (0.021) & 0.121 (0.015) & 0.426 (0.038) & 0.494 (0.049) & 0.375 (0.042) & 0.777 (0.036) \\
\bottomrule
\end{tabular}
\end{table}

>>>>

As described in the manuscript, the Martingale residuals $\{m_i\}_{i=1}^n$ and deviance residuals $\{d_i\}_{i=1}^n$ for the Cox model are given by
\begin{align*}
  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\beta) \\
  d_{i} &= \text{sign}(m_{i})\sqrt{-2(m_i + \delta_{i}\log(\delta_{i} - m_i))},
\end{align*}
where $H_0$ is the cumulative baseline hazard.  Our cross-validated versions of these residuals use
\begin{align*}
  m_{i}^{\text{cv}} &= \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}) \\
  d_{i}^{\text{cv}} &= \text{sign}(m_{i}^{\text{cv}})\sqrt{-2(m_i^{\text{cv}} + \delta_{i}\log(\delta_{i} - m_i^{\text{cv}}))},
\end{align*}
for $i$ in the test set $D_k$, with $\hat{\beta}^{-k}$ obtained from the training set $T_k$.  The sum of squared cross-validated deviance residuals, $\sum_{i}(d_{i}^{\text{cv}})^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression.  In the manuscript, this method is referred to as the \emph{cross-validated deviance residuals} approach.

@@

\begin{algorithm}
\SetAlgoLined
Use all observations' time and status to compute $\hat{H}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}$\;
\For{each regularization parameter $\lambda$}{
 \For{each fold $k = 1,2,...,K$}{
  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
  	\For{each obervation $i$ in the $k$th fold}{
  	Compute the Cross-validated Martingale Residual: $m^{\text{cv}}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(x_i\Tr\hat\bbeta^{-k})$\;
  	Compute the Deviance Residual: $ d^{\text{cv}}_{i} = \text{sgn}(m^{\text{cv}}_{i})\sqrt{-2(m^{\text{cv}}_{i} + \delta_{i}\log(\delta_{i} - m^{\text{cv}}_{i}))}$\;
  	}
 }
 Compute the sum of squares: $\sum_{i}({d}^{\text{cv}}_{i})^2$
 }
 \caption{\label{alg:na} Compute cross-validated deviance residuals based on the Nelson-Aalen estimator}
\end{algorithm}

@@

In the original proposal of the Martingale Residuals \citep{Therneau1990}, Breslow's estimator \citep{Breslow1974} was the prefered approach for estimating the cumulative baseline hazard. Under the assumption that there is no ties among the failure times, the Breslow's estimator is defined as:
\begin{equation} 
\hat{H}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}.
\end{equation}
Covariates of the fitted model are incorporated in the denominator of the summation.
Procedure~\ref{alg:brkp} includes an explicit description of this algorithm.
\begin{algorithm}
\SetAlgoLined
\For{each regularization parameter $\lambda$}{
 Fit the model using all observations' time and status \;
 Obtain linear predictor $\hat{\eta}$ from the fitted model\;
 \uIf{Use the Breslow's estimator}{
 $\hat{H}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}$;
 }\ElseIf{Use the Kalbfleisch and Prentice estimator}{
  $\hat{H}_0(t) = \sum_{t_{i} \leq t} \left\{1 -  \left(1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})} \right) ^{\exp(-\hat{\eta}_{i})}\right\}$\;
 }
 \For{each fold $k = 1,2,...,K$}{
  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
  	\For{each obervation $i$ in the $k$th fold}{
  	Compute the Cross-validated Martingale Residual: $m^{\text{cv}}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(x_i\Tr\hat\bbeta^{-k})$\;
  	Compute the Deviance Residual: $ d^{\text{cv}}_{i} = \text{sgn}(m^{\text{cv}}_{i})\sqrt{-2(m^{\text{cv}}_{i} + \delta_{i}\log(\delta_{i} - m^{\text{cv}}_{i}))}$\;
  	}
 }
 Compute the sum of squares: $\sum_{i}({d}^{\text{cv}}_{i})^2$
 }
 \caption{\label{alg:brkp} Compute cross-validated deviance residuals based on the Breslow estimator and the Kalbfleisch and Prentice estimator}
\end{algorithm}

@@

\begin{table}[ht]

@@

\begin{table}[ht]
\setlength{\tabcolsep}{3pt}
\caption{\label{Tab:wb2} Weibull Distribution, scale = 1, shape = 1.5}
\centering
\begin{tabular}[t]{lccccccccc}
\toprule
 Method & $\lambda$ & log(MSE Ratio) & Brier (25$\%$) & Brier(Median) & Brier(75$\%$) & KL (25$\%$)& KL (Median) & KL (75$\%$) & C Index\\
\midrule
&\multicolumn{9}{c}{\textbf{Scenario 1}: n = 150, p = 50, 5 non-zero features, weak signal}\\
\cline{2-10}
V VH  & 0.096 (0.026) & 1.502 (0.89) & 0.185 (0.006) & 0.244 (0.015) & 0.182 (0.014) & 0.558 (0.02) & 0.681 (0.032) & 0.545 (0.034) & 0.614 (0.03) \\
Standard  & 0.089 (0.03) & 1.479 (0.873) & 0.184 (0.007) & 0.243 (0.015) & 0.182 (0.013) & 0.557 (0.021) & 0.681 (0.032) & 0.544 (0.033) & 0.615 (0.031) \\
LinearPred  & 0.09 (0.027) & 1.482 (0.871) & 0.185 (0.007) & 0.243 (0.014) & 0.182 (0.014) & 0.558 (0.02) & 0.681 (0.031) & 0.544 (0.033) & 0.615 (0.03) \\
DevResid  & 0.098 (0.025) & 1.503 (0.869) & 0.185 (0.006) & 0.244 (0.015) & 0.182 (0.014) & 0.558 (0.02) & 0.681 (0.031) & 0.545 (0.033) & 0.615 (0.03) \\
\addlinespace
&\multicolumn{9}{c}{\textbf{Scenario 2}: n = 150, p = 50, 5 non-zero features, strong signal}\\
\cline{2-10}
 V VH  & 0.07 (0.011) & 1.733 (0.83) & 0.147 (0.009) & 0.176 (0.012) & 0.128 (0.009) & 0.453 (0.023) & 0.525 (0.028) & 0.395 (0.026) & 0.75 (0.013) \\ 
Standard  & 0.066 (0.011) & 1.693 (0.832) & 0.147 (0.009) & 0.176 (0.011) & 0.128 (0.009) & 0.452 (0.023) & 0.525 (0.028) & 0.395 (0.026) & 0.749 (0.014) \\
 LinearPred  & 0.067 (0.01) & 1.699 (0.834) & 0.147 (0.009) & 0.176 (0.011) & 0.128 (0.009) & 0.452 (0.023) & 0.525 (0.028) & 0.395 (0.026) & 0.749 (0.014) \\
 DevResid  & 0.096 (0.013) & 2.14 (0.799) & 0.151 (0.009) & 0.181 (0.012) & 0.13 (0.009) & 0.463 (0.022) & 0.539 (0.028) & 0.406 (0.026) & 0.751 (0.013) \\
\addlinespace
&\multicolumn{9}{c}{\textbf{Scenario 3}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
\cline{2-10}
V VH  & 0.126 (0.019) & 2.948 (0.442) & 0.188 (0.005) & 0.252 (0.011) & 0.191 (0.011) & 0.564 (0.016) & 0.696 (0.023) & 0.569 (0.026) & 0.585 (0.043) \\ 
Standard  & 0.103 (0.024) & 2.866 (0.465) & 0.184 (0.006) & 0.243 (0.014) & 0.183 (0.012) & 0.555 (0.019) & 0.68 (0.029) & 0.548 (0.03) & 0.6 (0.041) \\ 
LinearPred  & 0.104 (0.024) & 2.869 (0.464) & 0.185 (0.006) & 0.244 (0.014) & 0.183 (0.012) & 0.556 (0.019) & 0.68 (0.029) & 0.549 (0.03) & 0.6 (0.041) \\   
 DevResid  & 0.107 (0.023) & 2.88 (0.458) & 0.185 (0.006) & 0.244 (0.013) & 0.184 (0.012) & 0.556 (0.019) & 0.682 (0.028) & 0.55 (0.03) & 0.599 (0.041) \\
\addlinespace
&\multicolumn{9}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, strong signal}\\
\cline{2-10}
 V VH  & 0.082 (0.012) & 3.74 (0.554) & 0.145 (0.019) & 0.173 (0.029) & 0.127 (0.02) & 0.443 (0.05) & 0.519 (0.068) & 0.395 (0.058) & 0.766 (0.047) \\  
Standard  & 0.059 (0.009) & 3.414 (0.582) & 0.131 (0.015) & 0.154 (0.02) & 0.117 (0.014) & 0.405 (0.041) & 0.467 (0.05) & 0.367 (0.044) & 0.78 (0.03) \\   
LinearPred  & 0.06 (0.009) & 3.435 (0.583) & 0.132 (0.015) & 0.154 (0.02) & 0.117 (0.014) & 0.406 (0.041) & 0.468 (0.051) & 0.366 (0.043) & 0.78 (0.031) \\   
 DevResid  & 0.074 (0.008) & 3.66 (0.51) & 0.139 (0.014) & 0.163 (0.021) & 0.121 (0.015) & 0.426 (0.038) & 0.494 (0.049) & 0.375 (0.042) & 0.777 (0.036) \\
\bottomrule
\end{tabular}
\end{table}
